Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2019-07-19 paper/referee_reply2.txt

Reply: We have included discussion of all additional models we developed for purposes of the present paper.

3. Jackknife error bars

Referee: "Jackknife was one of the error estimation methods that I mentioned, but what about the scatter between different Monte Carlo realisations of the model. Presumably each realisation, with different M* and R_1/2 values for each subhalo, gives a single line on Figs 4-6, so that the entire model effectively produces a band even before jackknifing. How big is this scatter relative to that between different subregions for the same realisation? Or am I missing something and the line can only be calculated over many realisations? I don't think the role of the Monte Carlo realisations is explained completely clearly. Also, why not add the band to the green lines in Fig 6? It would be good to see explicitly that these are consistent with 0."

Reply: We have included jackknife error estimation the errors on the green curve in Figure 6. We have furthermore verified that for these clustering ratios, statistical errors associated with Monte Carlo realizations are approximately captured by the jackknife: when we repeat the error estimation by repopulating each of the 125 jackknife subsamples with a new realization, the error bars on the ratio plots are comparable to the more standard treatment in which mock population is done once. We consider this sufficient for the purpose of these error estimations, which is merely to guide the reader's eye, rather than to compute chi^2 in an MCMC analysis.

4. Section 6 discussion

### 2019-07-19 paper/referee_reply2.txt

We thank the referee for careful consideration of our revised manuscript. The second round of requested changes were rather minor, and so we apologize for our protracted reply.

In the intervening time we have pursued two closely related directions of research that we had originally thought might be included in our revisions: 1. Studying the IllustrisTNG predictions for size-dependent lensing, and 2. new SDSS measurements of size-dependentlensing. In the end, both extensions turned out to be more involved than we anticipated, and so we have decided to pursue these directions in separate papers.

We respond to each of the referee's second round of comments in the text below.

### 2019-08-06 paper/galsize_paper.tex
Implemented all of Andrey's comments on new submission except for the data table

For our SDSS galaxy samples, we calculate two-point clustering $\wproj$ using line-of-sight projection of $\pi_{\rm max}=20~\mpc$ using the {\tt correl} program in {\tt UniverseMachine} (Behroozi et al. 2017, in prep). For two samples of points distributed in redshift space, $\vec{x}_1$ and $\vec{x}_2,$ with $\vec{s}\equiv\vec{x}_2-\vec{x}_1$ and $\vec{l}\equiv0.5(\vec{x}_2+\vec{x}_1),$ the {\tt correl} code program counts the number of pairs of points within a line-of-sight distance $\pi\equiv(\vec{s}\cdot\vec{l})/\vert\vec{l}\vert$ and transverse separation $r_{\rm p}\equiv\sqrt{s^{2} - \pi^{2}}.$ When both sets of points are galaxy data, the resulting pair-counts are denoted by DD; when both sets are random points, RR; cross-counts between data and randoms are denoted by DR. Using the \citet{landy_szalay93} estimator ($DD - 2DR + RR$), we compute the redshift-space correlation function $\xi(\rproj, \pi)$, which is then integrated over $\vert\pi\vert < 20$ Mpc to compute the projected correlation function $\wproj(\rproj)$.  The code uses $10^6$ randoms drawn from the same mask region with uniform volume distribution to compute $DR$, and $RR$ is computed via Monte Carlo integration. Errors are estimated by jackknife resampling, giving us an estimate for the lower bound of samples with volumes $V_{\rm eff}\lesssim 0.3\,{\rm Gpc}^3.$

@@

In the $\rvir$-based size model, the weak dependence of $\rhalf$ on $\mstar$ for low-mass galaxies is due to the steep $\mstar-\mvir$ relation. For example, if $\mstar\propto\mvir^\alpha\propto\rvir^{3\alpha}$, then $\rhalf\propto \rvir\propto \mstar^{1/(3\alpha)}$. Given that $\alpha\sim 1-3$ is inferred for the $\mstar-\mvir$ relation at low masses \citep{kravtsov10,moster_etal13,behroozi13_smhm,kravtsov_etal14}, the $\rhalf-\mstar$ relation in this regime is expected to be shallow.
For galaxies of $\mstar\gtrsim 5\times 10^{10}\ M_\odot$, $\alpha\approx 0.3$ \citep[e.g.,][]{kravtsov_etal14}, and thus
the slope of the $\rhalf-\mstar$ relation is approximately linear. These trends of $\rhalf$ with $\mstar$ have a direct relation to the trends of surface brightness $\mu_\star=\mstar/(2\pi\rhalf^2)$ with stellar mass \citep[see, for example, the $\mu_\star-$dependent clustering measurements in][which are qualitatively similar to the measurements reported here]{li_etal06}.

@@

In other closely related work, efforts to constrain correlations between galaxy size and various halo properties at fixed {\em stellar mass} have been carried out using both weak lensing \citep[e.g.,][]{sonnenfeld_leauthaud18} and galaxy dynamics \citep[e.g.,][]{desmond_wechsler15,desmond_wechsler17}. Our results complement these previous studies: forward-modeling galaxies directly into simulations allows us to test hypotheses of correlations at fixed {\em halo mass.} Calculating fixed-$\mstar$ correlations from a Monte Carlo realization of the model is straightforward, and in our fiducial model, in which $\mean{\rhalf}{\rmpeak}=0.01\rmpeak,$ for Milky Way mass centrals of the same $\mstar$ there exists a significant positive correlation between present day halo mass $\mvir$ and galaxy size $\rhalf,$ in some tension with the finding based on internal dynamics that at fixed $\mstar,$ there is a weak anti-correlation between $\rhalf$ and halo mass \citep{desmond17a,desmond17b}. Alternative observational methods based on lensing and dynamics such as those mentioned above are a natural testbed for the conclusions drawn here based on two-point clustering. We now give an example of how such a future measurement could be made to directly and independently test these conclusions.

>>>>

For our SDSS galaxy samples, we calculate two-point clustering $\wproj$ using line-of-sight projection of $\pi_{\rm max}=20~\mpc$ using the {\tt correl} program in {\tt UniverseMachine} (Behroozi et al. 2017, in prep). For two samples of points distributed in redshift space, $\vec{x}_1$ and $\vec{x}_2,$ with $\vec{s}\equiv\vec{x}_2-\vec{x}_1$ and $\vec{l}\equiv0.5(\vec{x}_2+\vec{x}_1),$ the {\tt correl} code program counts the number of pairs of points within a line-of-sight distance $\pi\equiv(\vec{s}\cdot\vec{l})/\vert\vec{l}\vert$ and transverse separation $r_{\rm p}\equiv\sqrt{s^{2} - \pi^{2}}.$ When both sets of points are galaxy data, the resulting pair-counts are denoted by DD; when both sets are random points, RR; cross-counts between data and randoms are denoted by DR. Using the \citet{landy_szalay93} estimator ($DD - 2DR + RR$), we compute the redshift-space correlation function $\xi(\rproj, \pi)$. The function is then integrated over $\vert\pi\vert < 20$ Mpc to compute the projected correlation function $\wproj(\rproj)$.  The code uses $10^6$ randoms drawn from the same mask region with uniform volume distribution to compute $DR$, and $RR$ is computed via Monte Carlo integration. Errors are estimated by jackknife resampling, giving us an estimate for the lower bound of samples with volumes $V_{\rm eff}\lesssim 0.3\,{\rm Gpc}^3.$

@@

In the $\rvir$-based size model, the weak dependence of $\rhalf$ on $\mstar$ for low-mass galaxies is due to the steep $\mstar-\mvir$ relation. For example, if $\mstar\propto\mvir^\alpha\propto\rvir^{3\alpha}$, then $\rhalf\propto \rvir\propto \mstar^{1/(3\alpha)}$. Given that $\alpha\sim 1-3$ is inferred for the $\mstar-\mvir$ relation at low masses \citep{kravtsov10,moster_etal13,behroozi13_smhm,kravtsov_etal14}, the $\rhalf-\mstar$ relation in this regime is expected to be shallow.
For galaxies of $\mstar\gtrsim 5\times 10^{10}\ M_\odot$, $\alpha\approx 0.3$ \citep[e.g.,][]{kravtsov_etal18}, and thus
the slope of the $\rhalf-\mstar$ relation is approximately linear. These trends of $\rhalf$ with $\mstar$ have a direct relation to the trends of surface brightness $\mu_\star=\mstar/(2\pi\rhalf^2)$ with stellar mass \citep[see, for example, the $\mu_\star-$dependent clustering measurements in][which are qualitatively similar to the measurements reported here]{li_etal06}.

@@

In other closely related work, efforts to constrain correlations between galaxy size and various halo properties at fixed {\em stellar mass} have been carried out using both weak lensing \citep[e.g.,][]{sonnenfeld_leauthaud18} and galaxy dynamics \citep[e.g.,][]{desmond_wechsler15,desmond_wechsler17}. Our results complement these previous studies: forward-modeling galaxies directly in simulations allows us to test hypotheses of correlations at fixed {\em halo mass.} Calculating fixed-$\mstar$ correlations from a Monte Carlo realization of the model is straightforward, and in our fiducial model, in which $\mean{\rhalf}{\rmpeak}=0.01\rmpeak,$ for Milky Way mass centrals of the same $\mstar$ there exists a significant positive correlation between present day halo mass $\mvir$ and galaxy size $\rhalf,$ in some tension with the finding based on internal dynamics that at fixed $\mstar,$ there is a weak anti-correlation between $\rhalf$ and halo mass \citep{desmond17a,desmond17b}. Alternative observational methods based on lensing and dynamics such as those mentioned above are a natural testbed for the conclusions drawn here based on two-point clustering. We now give an example of how such a future measurement could be made to directly and independently test these conclusions.
